---
title: "R Session 1"
output: html_document
author: Sean Preusse
theme: united
highlight: zenburn
---

## Updates and Packages

#### Update R

if(!require(installr)) {install.packages("installr"); require(installr)} #load / install+load installr 

updateR()
 
This will start the updating process of your R installation.  It will check for newer versions, and if one is available, will guide you through the decisions you'd need to make.

#### Installing Packages

install.packages("packagename")

Examples of high use packages;
- ggplot2 # Data Visualisation
- ggvis # Interactive Visualtions
- xlsx # Load .xlsx files
- dplyr # Data Munging
- reshape2 # Data Munging, Melting
- rvest # Web Scraping
- lubridate # Dates
- corrplot # Correlation Plots


#### Loading Packages

```{r}
library(corrplot)
library(ggplot2)
library(ggvis)
```

## Creating Data

#### Strings
```{r}
Fruit <- c("Apples", "Bananas", "Oranges")
```

#### Values
```{r}
Quantity = c(4, 6, 9)
Cost <- c(.80, .60, .50)
```

#### Creating a table
```{r}
Shopping <- data.frame(Fruit, Quantity, Cost)
```

#### Add new calculation to table
```{r}
Shopping$Total.Cost <- Shopping$Quantity*Shopping$Cost
```

#### Get and Set working Directory

```{r}
getwd() 
# setwd("/Users/sppreus")
list.files()

#### Output a file
```{r}
write.table(Shopping, "Shopping.csv", sep=",")
rm(Shopping)
```

#### Loading file

```{r}
Shopping <- read.csv("Shopping.csv")
```

## Exploration
```{r}
Shopping
summary(Shopping)
# Total Shopping Cost
sum(Shopping$Total.Cost)
```

## Real Data Set Exploration - US Facts

We are using the preloaded dataset on US State Factiods

Dictionary
* Population: Estimate as of July 1975
* illiteracy: percentage of population
* Life Exp: Life expectancy in years
* Murder: Murder and non-negligent manslaughter rate per 100,000 population
* HS Grad: Percent high-school graduaates
* Frost: mean number of days with minimum temperature below freezing
* Area: Land area in square miles

## Describe Data

```{r}
us.factoids <- data.frame(state.x77)
str(us.factoids)
summary(us.factoids)
# Standard Deviation is important but missing, lets view this as well.
round(sapply(us.factoids, sd),2)
# Adding a Population Density
us.factoids$Density = us.factoids$Population * 1000 / us.factoids$Area
```

## Visualise and Plot Data
```{r}
#Histogram
qplot(Murder, data=us.factoids)
qplot(Murder, data=us.factoids, binwidth = 2)

#DensityPlot
qplot(Murder, geom="density", data=us.factoids)

#BoxPlot
us.factoids$State <- as.factor(rownames(us.factoids))
qplot(Murder, State, geom="boxplot", data=us.factoids)

#Scatter Plot
qplot(Murder, Population, geom='point', data=us.factoids, binwidth = 2)

#Non-Linear Regression with degrees of freedom
qplot(Murder, Population, geom=c("point","smooth"), data=us.factoids, binwidth = 2)
```

Using GGplot instead of qplot

```{r}
###BoxPlot
ggplot(data = us.factoids, aes(Murder, State)) + geom_boxplot()

### Lets make this easier to read
us.factoids$State <- factor(us.factoids$State, levels = us.factoids$State[order(us.factoids$Murder)])
ggplot(data = us.factoids, aes(Murder, State)) + geom_boxplot() + theme_bw()

#Scatter Plot
ggplot(data = us.factoids, aes(Murder, Population)) + geom_point() + geom_smooth() + theme_bw()
```

Interactive Chart Example

```{r}
us.factoids %>% ggvis(x= ~Murder) %>% layer_densities(adjust = input_slider(.1, 2, value =1, step =.1, label = "Bandwidth adjustment"))
```

## Multiple Regression example

#### Corrolation Matrix Table

```{r}
us.factoids$State <- NULL

us.factoids.cor <- cor(us.factoids, use="complete.obs")
round(us.factoids.cor,2)
```

#### Pairs Scatter Plot

```{r}
pairs(us.factoids)
```

#### Corrolation Matrix Plot

```{r}
library(corrplot)
?corrplot
corrplot(us.factoids.cor, order="AOE", method="pie", tl.pos="lt", type="upper",tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE,p.mat = 1-abs(us.factoids.cor), sig.level=0.70, insig = "blank")
```
